Stochastic Optimal Control of Single Neuron Spike Trains

### 2013-09-02 draft.tex
Camera-ready version for S+A approval + corrections

@@

\subsection{Parameter Regimes}
Different  parameter regimes can be envisioned given \cref{eq:X_evolution_uo},
depending on whether the noise intensity, $\b$, is
relatively high or low, and whether the external, uncontrollable bias current,
$\m$, induces spikes in the absence of noise or not. 
Spikes will occur in the absence of noise, if and only
if $ \m > \frac 1\tc,$ which is called the supra-threshold regime. When $\m
\leq 1/\tc$, the regime is called the sub-threshold regime.
In addition we will impose two values of $\b$, which we
will call high-noise and low-noise, respectively.
Example values for each parameter regime are given in \cref{tab:regimes}, and we
visualize a single path for each regime in \cref{fig:regime_path_examples}.
\begin{table}
\begin{tabular}{|l||{c}|{c}|}
\hline
\backslashbox{$\m$}{$\b$}
& $1.5$ & $0.3$ \\
\hline
$1.5 / \tc $ &Supra-threshold-High-noise & Supra-threshold-Low-noise \\
\hline
$0.1 / \tc$   &Sub-threshold-High-noise & Sub-threshold-Low-noise \\
\hline
\end{tabular}
\caption{Regime labels and example values. Note that for the numerical
experiments below, we use $\tc = 0.5$}
\label{tab:regimes}
\end{table}
%\usepackage{graphics} is needed for \includegraphics
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=0.99\textwidth]{Figs/PathSimulator/path_T=14_combined.pdf}
  \caption[labelInTOC]{Example Trajectories from \cref{eq:X_evolution_uo}
  using the parameter values from \cref{tab:regimes}. A)
  Supra-Threshold-High-Noise B) Supra-Threshold-Low-Noise C) Sub-Threshold-High-Noise D) Sub-Threshold-Low-Noise}
  \label{fig:regime_path_examples}
\end{center}
\end{figure}

@@

Now consider the differences between the individual regimes, i.e.\ the
effect of changing the bias and the noise intensity, $\m,\b$. All things being
equal, the effect of increasing the noise intensity, $\b$, is to lift the value
function, i.e.\ to make our objective worse, at the beginning of the interval
and to decrease it at the end. This is to be expected since we are attempting to
minimize the variance in the spike time and without noise there would be no
variance at all. However, at the end of the interval, noise only helps, since
now we are only interested in spiking as soon as possible, and a higher noise
intensity will tend to decrease the spike time on average. Increasing $\b$ also
has the effect of increasing the size of the boundary layer near $\xth$, where
the value function rises steeply. I.e.\ for small $\b$ that layer is small since
the risk of spiking early is only significant close to the threshold. For larger
$\b$ this layer increases, see panels A,C, with a thin layer, vs.\ B,D, with a
thicker layer, in \cref{fig:HJB_4regimes_value_control_cuts}. Similarly,
increasing the bias, $\mu$, tends to decrease the value function, especially at
the end since that has the unequivocal effect of preventing late spikes.
% \clearpage

\section{Open-loop Stochastic Control}
When the value of $X_t$ is unobservable, the best
one can do is to use the transition density to perform the optimization. Since the
transition density follows a (deterministic) PDE, we apply a Maximum
Principle for PDEs as a method of obtaining the optimal control,\cite{Borzi2012}.

\subsection{Fokker-Planck equation for the state density evolution}
We write the transition density of $X$, conditional on no
spikes having occurred as:
\begin{align*}
f(x,t) \intd{x} &= \Prob[X_t\in \intd{x} \,\big|\, X_0 = 0,\, X_{s} < \xth \quad
\forall s < t]
\end{align*}
Then $f$ satisfies a Fokker-Planck equation with absorbing boundaries,
see Ch.\ 7 in \cite{Jacobs},
\begin{equation}
\begin{gathered}
\di_t f(x,t) =
				\frac{\b^2 }{2}\cdot \di^2_x f -
				\di_x \Big[ \left( \m + \a(t)- \frac{x}{\tc}\right)  \cdot f \Big]
\\
\\
\begin{array}{lll}
	&\textrm{st.\ }&
	\left\{ \begin{array}{lcl}
	 f(x,0) &=& \delta(x) \quad \textrm{delta function}
	\\
	(\m + \a(t)- \tfrac x \tc)f - \di_x \tfrac {\b^2}2 f] \big|_{x=\xmin} &\equiv&
	0 \quad \textrm{reflecting BCs at some } \xmin
	\\
	f(x,t) |_{x=\xth} &\equiv& 0 \quad \textrm{absorbing BCs at } \xth
\end{array} \right.
\end{array}
\label{eq:FP_pde_OU_PDF}
\end{gathered}
\end{equation}
In theory, $\xmin = -\infty$, but in the numerics below we will need to
truncate it to some finite value, exactly as in the HJB equation,
\cref{eq:OC_LS_HJB_full}.
Note that we write $\a(t)$ here instead of $\a(x,t)$ since we cannot use the
value of $X_t$ in choosing the optimal control.

@@

\subsection{Restating the objective in terms of the transition density}
Let us recall our objective, \cref{eq:OC_LS_variance}:
$$
J[\a(\cdot)] = \Exp\left[
(\ts - \T \big)^2
+
\e \int_0^\ts  \a^2(s) \intd{s}
\right].
$$
Rewriting this in terms of the transition density, $f$, reads:
\begin{align}
J[\a(\cdot)] =&
\int_\xmin^\xth \Ttwo(x) \cdot f(x,\T) \intd{x}
\notag
\\
&+ \int_0^\T \phi(\xth, t) (t-\T)^2 \intd{t}
\label{eq:OC_LS_variance_density}
\\
&+  \e \int_0^\ts  \a^2(t)  \int_\xmin^\xth f(x,t) \intd{x} \intd{t}
\notag
\end{align}
Let us explain in more detail each term on the RHS in
\cref{eq:OC_LS_variance_density}.

@@

Finally, we are ready to calculate the open-loop stochastic optimal control for
the four parameter regimes. The calculated optimal controls for each regime are
in \cref{fig:FBK_Regimes_cs}. Recall that the optimal control, $\a^*(t)$, is
open-loop and thus it is only a function of time.
% \usepackage{graphics} is needed for \includegraphics
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/FP_Adjoint/Regimes_cs.pdf}
  \caption[labelInTOC]{The deterministic optimal controls for each parameter
  regime as functions of time, $t \in [0, \T]$.
  The desired spike time is set to $\T=1.5$, the energy penalty, $\eps
  = .001$ and the bounds are $\a \in [-2,2]$.
  A) Supra-Threshold-Low-Noise B) Supra-Threshold-High-Noise
C) Sub-Threshold-Low-Noise   D) Sub-Threshold-High-Noise  }
  \label{fig:FBK_Regimes_cs}
\end{center}
\end{figure}

@@

In addition to the optimal controllers, we also show the behaviour of
another control law - perhaps the most naive one - which is obtained by ignoring
the noise and the energy penalty. That is we find a {\sl constant} value of $\a$
which satisfies the desired boundary conditions, $x(0) = .0, x(\T) = \xth$. This naive
controller will be called 'Deterministic', since it assumes deterministic
dynamics in $X_t$, i.e.\ $\b = 0$.

@@

We set $\e = .001$ and $\amax = 2.0$. As such the energy cost is of secondary
importance and the paramount effect on the objective is the difference $(\ts -
\T)$, which we are trying to minimize.

@@

For illustration sake, we also visualize some trajectories from the
Sub-Threshold-High-Noise regime in \cref{fig:control_trajectories_examples}.

\begin{table}[h]
\begin{center}
\subfloat[Supra-Threshold High-Noise]{
\input{Figs/ControlSimulator/MeanSquaredErrors__SUPT_HN.txt}}\\
\subfloat[Supra-Threshold Low-Noise]{
   \input{Figs/ControlSimulator/MeanSquaredErrors__SUPT_ln.txt}}\\
 \subfloat[Sub-Threshold High-Noise]{
 \input{Figs/ControlSimulator/MeanSquaredErrors__subt_HN.txt}}\\
 \subfloat[Sub-Threshold Low-Noise]{
 \input{Figs/ControlSimulator/MeanSquaredErrors__subt_ln.txt}}\\
%  \begin{tabular}{lcc}
% Control Law &  Squared Error &  Squared Error \\
%  & Empirical & Theoretical\\% \hline
% Deterministic &  0.621 & - \\
% Open-Loop Stochastic & 0.356 & 0.382\\
% Feedback Stochastic &  0.288& 0.285\\
% \hline % \end{tabular} -
\caption{TODO(ALEX): THIS TABLE IS MOSTLY A SANITY CHECK TO SEE THAT THE CODE
WORKS - CONSIDER OMITTTING IN DRAFT.
Realized and theoretical performance of the different control laws. The
empirical performance is obtained using $N=10000$ sample paths. The theoretical
performance is found using the optimal value for $J$ for the open-loop
stochastic control and the value function $\v(x=0, t =0)$ for the closed-loop
stochastic control.}
\label{tab:realized_avg_errors_det_vs_openloop_vs_stoch}
\end{center}
\end{table}

@@

\begin{figure}[h]
\begin{center}
\subfloat[]
{
\label{fig:controlled_traj_ex1}
\includegraphics[width=0.33\textwidth]
{Figs/ControlSimulator/SubTHighNoise_Traj0.pdf}
}
\subfloat[ ]
{
\label{fig:controlled_traj_ex2}
\includegraphics[width=0.33\textwidth]
{Figs/ControlSimulator/SubTHighNoise_Traj4.pdf}
}
\subfloat[ ]
{
\label{fig:controlled_traj_ex3}
\includegraphics[width=0.33\textwidth]
{Figs/ControlSimulator/SubTHighNoise_Traj7.pdf}
}
\caption[]{Examples for the controlled trajectories using the
deterministic, open-loop stochastic and closed-loop stochastic control
approaches. The black vertical line in the plots indicates the desired spike-time, $\T$.
For the parameter values are $\m, \tc, \b = [.2, .5,  1.5]$ (Sub-Threshold
High-Noise regime). $\T = 1.5, \eps = .001$. The three panels from left to
right are three different realizations of the model dynamics. On the upper
plots, we show the voltage evolution, $X_t$, on the lower plots we show the
applied control, $\a(t)$. Note that $\a(t)$ obtained from the
deterministic and open-loop controls are the same for all three samples.}
\label{fig:control_trajectories_examples}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\subfloat[]
{
\label{fig:controlled_traj_ex1}
\includegraphics[width=0.33\textwidth]
{Figs/ControlSimulator/HighEps_Traj5.pdf}
}
\subfloat[ ]
{
\label{fig:controlled_traj_ex2}
\includegraphics[width=0.33\textwidth]
{Figs/ControlSimulator/HighEps_Traj6.pdf}
}
\subfloat[ ]
{
\label{fig:controlled_traj_ex3}
\includegraphics[width=0.33\textwidth]
{Figs/ControlSimulator/HighEps_Traj0.pdf}
}
\caption[]{The effect of higher energy penalty, $\eps = 0.1$.
Examples for the controlled trajectories using the deterministic,
open-loop stochastic and closed-loop stochastic control approaches. The black vertical line in the plots indicates the desired
spike-time, $\T$. For the parameter values are $\m, \tc, \b = [.2, .5,  1.5]$ (Sub-Threshold
High-Noise regime). $\T = 1.5, \eps = .1$. The three panels from left to
right are three different realizations of the model dynamics. On the upper
plots, we show the voltage evolution, $X_t$, on the lower plots we show the
applied control, $\a(t)$. Note that the optimal control, $\a(t)$, obtained from
the deterministic and from the open-loop controls are the same across all three
samples.}
\label{fig:control_trajectories_examples_high_epsilon}
\end{center}
\end{figure}

\subsection{Multi-Spike Control}
We now turn to the ultimate goal of our analysis, the control of {\sl spike
trains}. In particular, we will generate $M=10$ realizations of $N=20$ spikes
each in attempt to meet a prescribed spike train. We will focus on two
regimes, the sub-threshold regime with either low or high noise.
The results
for parameters from the Sub-Threshold, Low-Noise regime are shown in
results in \cref{fig:targettrain_cl_lownoise,fig:targettrain_ol_lownoise}.
While, the results
for parameters from the Sub-Threshold, High-Noise regime are shown in
results in \cref{fig:targettrain_cl_highnoise,fig:targettrain_ol_highnoise}.
%\usepackage{graphics} is needed for \includegraphics
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% CRIT REGIME LOW-NOISe:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/TrainController/CRITLN_cl_trains_sim_10.pdf}
  \caption[ ]{How good the {\sl closed-loop} controller is at hitting a target
  spike train. Uses the parameters from the CRITICAL,Low-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_cl_critlownoise}
\end{center}
\end{figure}
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/TrainController/CRITLN_ol_trains_sim_10.pdf}
  \caption[ ]{How good the {\sl open-loop} controller is at hitting a target
  spike train. Uses the parameters from the CRITICAL,Low-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_ol_critlownoise}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% CRIT REGIME HIGH-NOISe:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/TrainController/CRITHN_cl_trains_sim_10.pdf}
  \caption[ ]{How good the {\sl closed-loop} controller is at hitting a target
  spike train. Uses the parameters from the CRITICAL, High-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_cl_critlownoise}
\end{center}
\end{figure}
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/TrainController/CRITHN_ol_trains_sim_10.pdf}
  \caption[ ]{How good the {\sl open-loop} controller is at hitting a target
  spike train. Uses the parameters from the CRITICAL, High-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_ol_critlownoise}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% CRIT REGIME LOW-NOISe (AHmadian):
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/TrainController/CRITLN_Ahmadian_cl_trains_sim_10.pdf}
  \caption[ ]{AHMADIAN COMPARES: How good the {\sl closed-loop} controller is at
  hitting a target spike train. Uses the parameters from the CRITICAL, Low-Noise
  regime. The target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_cl_critlownoise}
\end{center}
\end{figure}
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/TrainController/CRITLN_Ahmadian_ol_trains_sim_10.pdf}
  \caption[ ]{AHMADIAN COMPARES: How good the {\sl open-loop} controller is at
  hitting a target spike train. Uses the parameters from the CRITICAL, Low-Noise
  regime. The target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_ol_critlownoise}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% FIGS:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
\begin{center}
\subfloat[DO NOT SKIP MISSED]{
  \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTLN_cl_trains_sim_10.pdf}
  }\\
 \subfloat[SKIP MISSED]{
  \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTLN_cl_skip_trains_sim_10.pdf}
  }
  \caption[ ]{How good the {\sl closed-loop} controller is at hitting a target
  spike train. Uses the parameters from the Sub-Threshold,Low-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the

>>>>

\tableofcontents

@@

\subsection{Parameter Regimes}
Different  parameter regimes can be envisioned given \cref{eq:X_evolution_uo},
depending on whether the noise intensity, $\b$, is
relatively high or low, and whether the external, uncontrollable bias current,
$\m$, induces spikes in the absence of noise or not. 
Spikes will occur in the absence of noise, if and only
if $ \m > \frac 1\tc,$ which is called the supra-threshold regime. When $\m
\leq 1/\tc$, the regime is called the sub-threshold regime.
In addition we will impose two values of $\b$, which we
will call high-noise and low-noise, respectively.
Example values for each parameter regime are given in \cref{tab:regimes}, and we
visualize a single path for each regime in \cref{fig:regime_path_examples}.
\begin{table}
\begin{tabular}{|l||{c}|{c}|}
\hline
\backslashbox{$\m$}{$\b$}
& $1.5$ & $0.3$ \\
\hline
$1.5 / \tc $ &Supra-threshold-High-noise & Supra-threshold-Low-noise \\
\hline
$0.1 / \tc$   &Sub-threshold-High-noise & Sub-threshold-Low-noise \\
\hline
\end{tabular}
\caption{Regime labels and example values. Note that for the numerical
experiments below, we use $\tc = 0.5$}
\label{tab:regimes}
\end{table}
%\usepackage{graphics} is needed for \includegraphics
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=0.99\textwidth]{Figs/PathSimulator/path_T=14_combined.pdf}
  \caption[labelInTOC]{Example Trajectories from \cref{eq:X_evolution_uo}
  using the parameter values from \cref{tab:regimes}. 
  A) Supra-Threshold-Low-Noise 
  B)Supra-Threshold-High-Noise 
  C) Sub-Threshold-Low-Noise
  D) Sub-Threshold-High-Noise 
  }
  \label{fig:regime_path_examples}
\end{center}
\end{figure}

@@

Now consider the differences between the individual regimes, i.e.\ the
effect of changing the bias and the noise intensity, $\m,\b$. All things being
equal, the effect of increasing the noise intensity, $\b$, is to lift the value
function, i.e.\ to make our objective worse, at the beginning of the interval
and to decrease it at the end. This is to be expected since we are attempting to
minimize the variance in the spike time and without noise there would be no
variance at all. However, at the end of the interval, noise only helps, since
now we are only interested in spiking as soon as possible, and a higher noise
intensity will tend to decrease the spike time on average. Increasing $\b$ also
has the effect of increasing the size of the boundary layer near $\xth$, where
the value function rises steeply. I.e.\ for small $\b$ that layer is small since
the risk of spiking early is only significant close to the threshold. For larger
$\b$ this layer increases, see panels A,C, with a thin layer, vs.\ B,D, with a
thicker layer, in \cref{fig:HJB_4regimes_value_control_cuts}. Similarly,
increasing the bias, $\mu$, tends to decrease the value function, especially at
the end since that has the unequivocal effect of preventing late spikes.

\section{Open-loop Stochastic Control}
When the value of $X_t$ is unobservable, the transition density can be used to
perform the optimization. Since the transition density follows a deterministic
PDE, we apply a Maximum Principle for PDEs as a method of obtaining the optimal
control, see \cite{Borzi2012} for details on optimization with PDEs.

\subsection{Fokker-Planck equation for the state density evolution}
We write the transition density of $X$, conditional on no
spikes having occurred as:
\begin{align*}
f(x,t) \intd{x} &= \Prob[X_t\in \intd{x} \,\big|\, X_0 = 0,\, X_{s} < \xth \quad
\forall s < t]
\end{align*}
Then $f$ satisfies a Fokker-Planck equation with absorbing boundaries,
see Ch.\ 7 in \cite{Jacobs},
\begin{equation}
\begin{gathered}
\di_t f(x,t) =
				\frac{\b^2 }{2}\cdot \di^2_x f -
				\di_x \Big[ \left( \m + \a(t)- \frac{x}{\tc}\right)  \cdot f \Big]
\\
\\
\begin{array}{lll}
	&\textrm{st.\ }&
	\left\{ \begin{array}{lcl}
	 f(x,0) &=& \delta(x) \quad \textrm{delta function}
	\\
	(\m + \a(t)- \tfrac x \tc)f - \di_x \tfrac {\b^2}2 f] \big|_{x=\xmin} &\equiv&
	0 \quad \textrm{reflecting BCs at some } \xmin
	\\
	f(x,t) |_{x=\xth} &\equiv& 0 \quad \textrm{absorbing BCs at } \xth
\end{array} \right.
\end{array}
\label{eq:FP_pde_OU_PDF}
\end{gathered}
\end{equation}
In theory, $\xmin = -\infty$, but in the numerics below we will need to
truncate it to some finite value, exactly as in the HJB equation,
\cref{eq:OC_LS_HJB_full}.
Note that we write $\a(t)$ here instead of $\a(x,t)$ since we cannot use the
value of $X_t$.

@@

\subsection{Restating the objective in terms of the transition density}
Let us recall our objective, \cref{eq:OC_LS_variance}:
$$
J[\a(\cdot)] = \Exp\left[
(\ts - \T \big)^2
+
\e \int_0^\ts  \a^2(s) \intd{s}
\right].
$$
Rewriting this in terms of the transition density, $f$, reads:
\begin{align}
J[\a(\cdot)] =&
\int_\xmin^\xth \Ttwo(x) \cdot f(x,\T) \intd{x}
\notag
\\
&+ \int_0^\T \phi(\xth, t) (t-\T)^2 \intd{t}
\label{eq:OC_LS_variance_density}
\\
&+  \e \int_0^\ts  \a^2(t)  \int_\xmin^\xth f(x,t) \intd{x} \intd{t}
\notag
\end{align}
Let us explain in more detail each term on the right-hand side in
\cref{eq:OC_LS_variance_density}.

@@

Finally, we are ready to calculate the open-loop stochastic optimal control for
the four parameter regimes. The calculated optimal controls for each regime are
in \cref{fig:FBK_Regimes_cs}. Recall that the optimal control, $\a^*(t)$, is
open-loop and thus it is only a function of time.
% \usepackage{graphics} is needed for \includegraphics
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/FP_Adjoint/Regimes_cs.pdf}
  \caption[labelInTOC]{The deterministic optimal controls for each parameter
  regime as functions of time, $t \in [0, \T]$.
  The desired spike time is set to $\T=1.5$, the energy penalty, $\eps
  = 0.001$ and the bounds are $\a \in [-2,2]$.
  A) Supra-Threshold-Low-Noise B) Supra-Threshold-High-Noise
C) Sub-Threshold-Low-Noise   D) Sub-Threshold-High-Noise  }
  \label{fig:FBK_Regimes_cs}
\end{center}
\end{figure}

@@

In addition to the optimal controllers, we also show the behaviour of
another control law - perhaps the most naive one - which is obtained by ignoring
the noise and the energy penalty. That is we find a {\sl constant} value of $\a$
which satisfies the desired boundary conditions, $x(0) = 0, x(\T) = \xth$. This naive
controller will be called 'Deterministic', since it assumes deterministic
dynamics in $X_t$, i.e.\ $\b = 0$.

@@

We set $\e = 0.001$ and $\amax = 2.0$. As such the energy cost is of secondary
importance and the paramount effect on the objective is the difference $(\ts -
\T)$, which we are trying to minimize.

@@

For illustration sake, we also visualize some trajectories from the
Sub-Threshold-High-Noise regime in \cref{fig:control_trajectories_examples},
see panels A),C),E).

\begin{table}[h]  
\begin{center}  
\subfloat[Supra-Threshold Low-Noise]{
   \input{Figs/ControlSimulator/MeanSquaredErrors__SUPT_ln.txt}}\\ 
\subfloat[Supra-Threshold High-Noise]{   
\input{Figs/ControlSimulator/MeanSquaredErrors__SUPT_HN.txt}}\\
 \subfloat[Sub-Threshold Low-Noise]{
 \input{Figs/ControlSimulator/MeanSquaredErrors__subt_ln.txt}}\\ 
 \subfloat[Sub-Threshold High-Noise]{
 \input{Figs/ControlSimulator/MeanSquaredErrors__subt_HN.txt}}\\
%  \begin{tabular}{lcc}
% Control Law &  Squared Error &  Squared Error \\
%  & Empirical & Theoretical\\% \hline 
% Deterministic &  0.621 & - \\
% Open-Loop Stochastic & 0.356 & 0.382\\
% Feedback Stochastic &  0.288& 0.285\\
% \hline % \end{tabular} - 
\caption{
Realized and theoretical performance of the different control laws. The
empirical performance is obtained using $N=10000$ sample paths. The theoretical
performance is found using the optimal value for $J$ for the open-loop
stochastic control and the value function $\v(x=0, t =0)$ for the closed-loop
stochastic control.}
\label{tab:realized_avg_errors_det_vs_openloop_vs_stoch} 
\end{center}
\end{table}

@@

% \begin{figure}[h]
% \begin{center}
% \subfloat[]
% {
% \label{fig:controlled_traj_ex1}
% \includegraphics[width=0.33\textwidth]
% {Figs/ControlSimulator/SubTHighNoise_Traj0.pdf}
% }
% \subfloat[ ]
% {
% \label{fig:controlled_traj_ex2}
% \includegraphics[width=0.33\textwidth]
% {Figs/ControlSimulator/SubTHighNoise_Traj4.pdf}
% }
% \subfloat[ ]
% {
% \label{fig:controlled_traj_ex3}
% \includegraphics[width=0.33\textwidth]
% {Figs/ControlSimulator/SubTHighNoise_Traj7.pdf}
% }
% \caption[]{Examples for the controlled trajectories using the
% deterministic, open-loop stochastic and closed-loop stochastic control
% approaches. The black vertical line in the plots indicates the desired spike-time, $\T$.
% For the parameter values are $\m, \tc, \b = [.2, .5,  1.5]$ (Sub-Threshold
% High-Noise regime). $\T = 1.5, \eps = 0.001$. The three panels from left to
% right are three different realizations of the model dynamics. On the upper
% plots, we show the voltage evolution, $X_t$, on the lower plots we show the
% applied control, $\a(t)$. Note that $\a(t)$ obtained from the
% deterministic and open-loop controls are the same for all three samples.}
% \label{fig:control_trajectories_examples}
% \end{center}
% \end{figure}

\subsection{Multi-Spike Control}
We now turn to the ultimate goal of our analysis, the control of {\sl spike
trains}. In particular, we will generate $M=10$ realizations of $N=20$ spikes
each in an attempt to meet a prescribed spike train. We will focus on two
regimes, the sub-threshold regime with either low or high noise.
The results
for parameters from the Sub-Threshold, Low-Noise regime are shown in
results in \cref{fig:targettrain_cl_lownoise,fig:targettrain_ol_lownoise}.
While, the results
for parameters from the Sub-Threshold, High-Noise regime are shown in
results in
\cref{fig:targettrain_cl_SUPT_HN,fig:targettrain_cl_SUPT_ln,fig:targettrain_ol_SUPT_ln,fig:targettrain_ol_SUPT_HN}.
%\usepackage{graphics} is needed for \includegraphics
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% SUPER-THRESHOLD REGIME LOW-NOISe:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/TrainController/SUPT_ln_cl_trains_sim_10.pdf}
  \caption[ ]{How good the {\sl closed-loop} controller is at hitting a target
  spike train. Uses the parameters from the Super-Threshold,Low-Noise regime.
  The target spike train was generated using the model itself, with
  parameters from the Super-Threshold, High-Noise regime, without an applied
  control, $\a=0$.
  There are $N=16$ spikes and $M=10$ realizations of the control strategy} 
  \label{fig:targettrain_cl_SUPT_ln}
\end{center}
\end{figure}  
\begin{figure}[htp] 
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/TrainController/SUPT_ln_ol_trains_sim_10.pdf}
  \caption[ ]{How good the {\sl open-loop} controller is at hitting a target
  spike train. Uses the parameters from the Super-Threshold,Low-Noise regime.
  The target spike train was generated using the model itself, with
  parameters from the Super-Threshold, High-Noise regime, without an applied
  control, $\a=0$.
  There are $N=16$ spikes and $M=10$ realizations of the control strategy}
  \label{fig:targettrain_ol_SUPT_ln}
\end{center} 
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% SUPER-THRESHOLD REGIME HIGH-NOISe:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/TrainController/SUPT_HN_cl_trains_sim_10.pdf}
  \caption[ ]{How good the {\sl closed-loop} controller is at hitting a target
  spike train. Uses the parameters from the Super-Threshold,High-Noise regime.
  The target spike train was generated using the model itself, with
  parameters from the Super-Threshold, High-Noise regime, without an applied
  control, $\a=0$.
  There are $N=16$ spikes and $M=10$ realizations of the control strategy}
  \label{fig:targettrain_cl_SUPT_HN}
\end{center} 
\end{figure}
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/TrainController/SUPT_HN_ol_trains_sim_10.pdf}
  \caption[ ]{How good the {\sl open-loop} controller is at hitting a target
  spike train. Uses the parameters from the Super-Threshold,High-Noise regime.
  The target spike train was generated using the model itself, with
  parameters from the Super-Threshold, High-Noise regime, without an applied
  control, $\a=0$.
  There are $N=16$ spikes and $M=10$ realizations of the control strategy}
  \label{fig:targettrain_ol_SUPT_HN}    
\end{center}
\end{figure}

\clearpage

\section{Miscleaneous Analysis???}
TODO: WE DON''T KNOW WHAT TO CALL THIS SECTION!
\subsection{Effect of energy penalty}
\label{sec:effect_of_eps}
So far we have assumed a low value for the energy penalty parameter,
$\eps$, in the objective defined in \cref{eq:OC_LS_variance_energy}. Our
approach has been that we are most concerned with the accurate spiking and that
penalizing energy is rather intended to regularize the control and avoid
excessive chattering of the control between its extreme values, than as a goal
in minimizing energy in its own right. Now, we take some time to explore the
effect of a higher $\eps$ on the optimal controls. Intuitively, a
higher value of $\eps$ will tend to bring the optimal control, $\a^*$, closer
to zero. We make a comparison for the four regimes for both the
closed-loop control and the open-loop control in
\cref{fig:HJB_4regimes_control_different_eps,fig:FBK_Regimes_cs_different_es}.
Indeed, for both the closed-loop and open-loop optimal controls, increasing
$\eps$ tends to reduce the absolute value of $\a^*(t)$. In particular, for the
closed-loop control it tends to broaden the area of transition for
decreasing $x$ when the control swings from its minimal, i.e.\ inhibitory value,
to its maximal, i.e.\ excitatory value. Similarly for the open-loop control,
instead of banging from its minimal bound at the beginning of the interval to
its maximal bound at the end, the optimal control for $\e = 0.1$ tends to have a
more mild transition from a slightly inhibitory to slightly excitatory
values.
